---
title: "Car prices"
author: "Aleix, David"
output:
    pdf_document:
        latex_engine: lualatex
---

```{r, libraries, include=FALSE}
library(tidyverse)
library(glue)
```

# Data loading

Load and bind datasets into single one. Adding manufacturer column.

```{r, message = FALSE}
df <- map2(# map through file and manufacturer names and read dataframes
           c("audi",  "bmw", "merc",     "vw"),         # filename
           c("Audi",  "BMW", "Mercedes", "Volkswagen"), # manufacturer
           function(filename, manufacturer) {
               read_csv(glue("./data/{filename}.csv"),
                        col_types = "fiififidd") %>%
               mutate(manufacturer = as_factor(manufacturer)) # add column
           }) %>%
    reduce(~ bind_rows(.x, .y)) # Bind rows into single dataframe
```

```{r, sample_size, include=FALSE}
glimpse(df)
```

Get a sample of `r (sample_size <- 1000)`:

```{r, sampleData}
set.seed(19990428)
df <- df %>%
    slice_sample(n = sample_size)
```

Add manufacturer to model column:

```{r, models}
df <- df %>%
    mutate(model = as_factor(paste0(manufacturer, " - ", model)))
```

Convert year and engineSize to factor. We do it now so they have numeric
ordering.

```{r, factors}
df <- df %>%
    mutate(year = as_factor(year),
           engineSize = as_factor(engineSize),
    )
```

Check for NA.

```{r, na}
df %>%
    summarise_all(~ sum(is.na(.))) %>%
    glimpse()
```

```{r, load_sample}
sample <- read_rds("./data/sample_clean_dataset.rds")
summary(sample)
summary(df)
glimpse(sample)
glimpse(df)
```

```{r, save, run=FALSE}
write_rds(df, "./data/dataset.rds")
```
