```{r, libraries, include=FALSE}
library(tidyverse)
library(glue)
library(chemometrics)
library(missForest)
library(cowplot)
library(doParallel)
```

# Data preparation

## Reading and initial preprocessing

Load and bind datasets into single one. Adding manufacturer column.

```{r read_data, echo = TRUE, message = FALSE}
df <- map2(# map through file and manufacturer names and read dataframes
    c("audi", "bmw", "merc", "vw"), # filename
    c("Audi", "BMW", "Mercedes", "Volkswagen"), # manufacturer
    function(filename, manufacturer) {
        read_csv(glue("./data/{filename}.csv"),
            col_types = "fiififidd"
        ) %>%
            mutate(manufacturer = as_factor(manufacturer)) # add column
    }
) %>%
    reduce(~ bind_rows(.x, .y)) # Bind rows into single dataframe
```

We took a sample of 5000 elements.

```{r sample_data, echo = TRUE}
set.seed(19990428)
df <- df %>%
    slice_sample(n = 5000)
```

## Univariate Descriptive analysis

We have `r ncol(df)` variables:

```{=latex }
\begin{multicols}{3}
{
```
```{r prep-remove-cols, results='asis'}
cat(paste0("- `", colnames(df), "`", sep = "\n"))
```
```{=latex }
}
\end{multicols}
```

From these variables, `r df %>% select(where(is.numeric)) %>% ncol()` are
numeric. In \cref{fig:num_summary} we see the distribution of these
numeric variables with boxplots.

```{r num_summary, fig.cap = "Boxplots of numeric variables in the dataset", fig.height = 5}
numeric_boxplots <- df %>%
    select(where(is.numeric)) %>%
    colnames() %>%
    map(~ ggplot(df, aes_string(.)) +
        geom_boxplot() +
        theme(axis.title.y = element_blank(),
              axis.text.y = element_blank(),
              axis.ticks.y = element_blank())
    )

plot_grid(plotlist = numeric_boxplots, ncol = 1)
```

The variables year corresponds to a qualitative concept and thus it should be
treated as a factor, to complement this change we add a new variable `age`
which corresponds to the age of the car. Given that the dataset is from
2020, we compute `age = 2020 - year`. This variable is numeric. Additionally,
we add auxiliary variables to the numeric ones that discretize them into
intervals. To simplify the intervals, the `price` and `mileage` values
in the auxiliary variables where divided by 1000.

The variable `engineSize` was converted to a factor since it can be argued that
it is a qualitative concept and there are a finite number of engine sizes in
the dataset.

We added `manufacturer` to the model column just in case there where models with
the same name from different manufacturers.

```{r aux_variables, echo = TRUE}
df <- df %>% mutate(
    model = as_factor(paste0(manufacturer, " - ", model)),
    age = 2020 - year,
    aux_price = cut_number(price / 1000, 4),
    aux_mileage = cut_number(mileage / 1000, 4),
    aux_mpg = cut_number(mpg, 4),
    aux_tax = cut_number(tax, 2),
    aux_age = cut_number(age, 4),
    year = as_factor(year),
    engineSize = as_factor(engineSize)
)
```

## Summary

\Cref{tab:numeric_summary} shows a summary of the numeric variables. Likewise,
\cref{tab:factor_summary} shows a summary of the categorical variables excluding
`model` and `engineSize`.


```{r numeric_summary, echo = FALSE}
df %>%
    select(where(is.numeric)) %>%
    papeR::summarize_numeric() %>%
    kable(caption = "Summary of numeric variables", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE)
```

```{r factor_summary, echo = FALSE}
df %>%
    select(where(is.factor), -c(year, model, engineSize), -starts_with("aux_")) %>%
    papeR::summarize_factor() %>%
    kable(caption = "Summary of categorical variables", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE) %>%
    footnote(general = "year, model and engineSize omitted")
```

```{r aux_summary, echo = FALSE}
df %>%
    select(starts_with("aux_")) %>%
    papeR::summarize_factor() %>%
    kable(caption = "Summary of auxiliary factor variables", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE)
```

There are `r nlevels(df$model)` different models
and `r nlevels(df$engineSize)` different engine sizes. \Cref{fig:engine_size}
shows the distribution of `engineSize`. In \cref{fig:car_models} we show the
15 most common car models in our sample.

```{r engine_size, fig.cap = "Distribution of engine sizes in the sample"}
df %>%
    ggplot(aes(engineSize)) + geom_bar() +
    labs(x = "Engine Size (L)", y = "n")
```

```{r car_models, fig.cap = "Most popular car models"}
df %>%
    group_by(model) %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    head(n = 15) %>%
    ggplot(aes(n, reorder(model, n))) +
    geom_bar(stat = "identity") +
    labs(y = "Model", x = "Count")
```

If we count the number of `NA` values per row, we find that there are no
explicit `NA` in the sample, as shown in \cref{tab:na}:

```{r na, echo = FALSE}
missings <- df %>%
    select(-starts_with("aux_"), -age) %>%
    summarise_all(~ sum(is.na(.))) %>%
    pivot_longer(everything(),
        names_to = "Variable",
        values_to = "Missing"
    )

zeros <- df %>%
    select(-starts_with("aux_"), -age) %>%
    summarise_all(~ sum(. == 0)) %>%
    pivot_longer(everything(),
        names_to = "Variable",
        values_to = "Zeros"
    )

left_join(missings, zeros, by = "Variable") %>%
    kable(caption = "Number of missing and zero values per row", booktabs = TRUE) %>%
    kable_styling(
        latex_options = c("HOLD_position"),
        full_width = FALSE
    )
```

## Outliers

### Severe outliers

To find severe outliers, for each numeric variable, we compute the IQR and check
which values are outside the range `(Q1 - 3*IQR, Q3 + 3*IQR)`.

```{r severe_outliers}
is_severe_outlier <- function(x) {
    tmp <- 3 * IQR(x, na.rm = TRUE)
    a <- quantile(x, 0.25, na.rm = TRUE) - tmp
    b <- quantile(x, 0.75, na.rm = TRUE) + tmp
    !dplyr::between(x, a, b)
}

# Create logical columns labelling values as outliers
df_outliers <- df %>%
    mutate(across(where(is.numeric),
        is_severe_outlier,
        .names = "sout_{col}")) %>%
    mutate(n_outliers = rowSums(across(starts_with("sout_"))))
```


\Cref{tab:outliers_ind} shows how many individuals have 0, 1, 2 or 3 outliers (
there are no individuals with more than 3 severe outliers).

```{r outliers_ind}
df_outliers %>% group_by(n_outliers) %>% summarise(count = n()) %>%
    kable(caption = "Number of severe outliers per individuals", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE)
```

The cars with 3 outliers are shown in \cref{tab:outliers_3} and have outliers
in `tax`, `mileage` and `age`.

```{r outliers_3}
df_outliers %>%
    select(model, n_outliers, year, mileage, tax, age) %>%
    arrange(desc(n_outliers)) %>%
    filter(n_outliers > 2) %>%
    relocate(n_outliers) %>%
    kable(caption = "Number of severe outliers per variable", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE)
```


```{r soutliers_summary}
df_outliers %>%
    summarise(across(starts_with("sout_"), sum)) %>%
    rename_with(~ gsub("sout_", "", .)) %>%
    kable(caption = "Number of severe outliers per variable", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE)
```

In \cref{tab:soutliers_summary} we can see that `tax` has a very high number of
severe outliers. If we plot the density function for the variable as shown in
\cref{fig:density_tax}, we can see that most of the values are around
`r median(df$tax)` and all the other peaks are labeled as severe outliers
since the IQR is `r IQR(df$tax)`. There is clearly a group of cars which pay
lower taxes, this may be correlated with other variables such as `fuelType`
of `engineSize`.

```{r density_tax, fig.cap = "Tax density plot with IQR"}
tax_mean <- mean(df$tax)
tax_iqr <- IQR(df$tax)
a <- quantile(df$tax, 1 / 4) - tax_iqr
b <- quantile(df$tax, 3 / 4) + tax_iqr

ggplot(df, aes(tax)) +
    geom_density() +
    geom_vline(aes(xintercept = tax_mean, color = "IQR")) +
    geom_vline(aes(xintercept = a, color = "IQR"), linetype = "dashed") +
    geom_vline(aes(xintercept = b, color = "IQR"), linetype = "dashed") +
    theme(legend.title = element_blank())
```

### Multivariate outliers

```{r moutliers}
res_outliers <- df %>%
    select(where(is.numeric), -year, -tax) %>%
    Moutlier(quantile = 0.999, plot = FALSE)

df_outliers <- df_outliers %>%
    add_column(
        moutlier_md = res_outliers$md,
        moutlier = res_outliers$md > res_outliers$cutoff
    )
```

To detect multivariate outliers, we use  `Moutlier` form the `chemometrics`
package. We found `r sum(df_outliers$moutlier)` multivariate outliers.
\Cref{tab:moutliers_list} shows a list of the 10 individuals with biggest Mahalanobis
distance.

```{r moutliers_list}
df_outliers %>%
    filter(moutlier) %>%
    select(model, age, price, mileage, tax, mpg, fuelType, engineSize,
           transmission, moutlier_md) %>%
    arrange(desc(moutlier_md)) %>%
    head(n = 10) %>%
    kable(caption = "Top 10 multivariate outliers", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position", "scale_down"), full_width = FALSE)
```


## Errors and inconsistencies

There where only 3 electric cars in the original dataset before the sample, in
our sample, we have no electric cars, however there where cars with
`engineSize` 0.  As shown in \cref{tab:engineZero}. Since they where not
classified as `Other` we decided that this was erroneous data which should be
imputed.

```{r engineZero}
df %>%
    filter(engineSize == 0) %>%
    group_by(fuelType) %>%
    summarise(n = n()) %>%
    kable(caption = "Individuals with engineSize 0 by fuelType", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE)

df_outliers <- df_outliers %>%
    mutate(imputed = engineSize == 0)

df[df$engineSize == "0", "engineSize"] <- NA
```

```{r imputation, results='hide', cache = TRUE}
registerDoParallel(cores=detectCores())
df_imp <- df %>%
    select(-c(model, -starts_with("aux_"))) %>%
    as.data.frame() %>%
    missForest(parallelize = "forests")

oob_error <- df_imp$OOBerror

df <- df %>% mutate(engineSize = df_imp$ximp$engineSize)
```

```{r save, include = FALSE}
source("io_checksum.R")
write_rds_w_checksum(df, "./data/dataset.rds")
```
