```{r, libraries, include=FALSE}
library(tidyverse)
library(glue)
```

# Data preparation

## Reading and intial preprocessing

Load and bind datasets into single one. Adding manufacturer column.

```{r read_data, message = FALSE}
df <- map2(# map through file and manufacturer names and read dataframes
    c("audi", "bmw", "merc", "vw"), # filename
    c("Audi", "BMW", "Mercedes", "Volkswagen"), # manufacturer
    function(filename, manufacturer) {
        read_csv(glue("./data/{filename}.csv"),
            col_types = "fiififidd"
        ) %>%
            mutate(manufacturer = as_factor(manufacturer)) # add column
    }
) %>%
    reduce(~ bind_rows(.x, .y)) # Bind rows into single dataframe
```

Get a sample of 5000:

```{r sample_data}
set.seed(19990428)
df <- df %>%
    slice_sample(n = 5000)
```

Add `manufacturer` to `model`, convert `year` and `engineSize` to factors and add
auxiliary factor variables for the numeric ones:

```{r aux_variables}
df <- df %>% mutate(
    model = as_factor(paste0(manufacturer, " - ", model)),
    age = 2020 - year,
    aux_price = cut_number(price / 1000, 4),
    aux_mileage = cut_number(mileage / 1000, 4),
    aux_mpg = cut_number(mpg, 4),
    aux_tax = cut_number(tax, 2),
    aux_age = cut_number(age, 4),
    year = as_factor(year),
    engineSize = as_factor(engineSize)
)
```

## Summary

```{r numeric_summary, echo = FALSE}
df %>%
    select(where(is.numeric)) %>%
    papeR::summarize_numeric() %>%
    kable(caption = "Summary of numeric variables", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE)
```

```{r factor_summary, echo = FALSE}
df %>%
    select(where(is.factor), -c(year, model, engineSize), -starts_with("aux_")) %>%
    papeR::summarize_factor() %>%
    kable(caption = "Summary of numeric variables", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE) %>%
    footnote(general = "year, model and engineSize omitted")
```

```{r aux_summary, echo = FALSE}
df %>%
    select(starts_with("aux_")) %>%
    papeR::summarize_factor() %>%
    kable(caption = "Summary of auxiliary factor variables", booktabs = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position"), full_width = FALSE)
```

If we count the number of `NA` values per row, we find that there are no
explicit `NA` in the sample, as shown in \cref{tab:na}:

```{r na, echo = FALSE}
missings <- df %>%
    select(-starts_with("aux_"), -age) %>%
    summarise_all(~ sum(is.na(.))) %>%
    pivot_longer(everything(),
        names_to = "Variable",
        values_to = "Missing"
    )

zeros <- df %>%
    select(-starts_with("aux_"), -age) %>%
    summarise_all(~ sum(. == 0)) %>%
    pivot_longer(everything(),
        names_to = "Variable",
        values_to = "Zeros"
    )

left_join(missings, zeros, by = "Variable") %>%
    kable(caption = "Number of missing and zero values per row", booktabs = TRUE) %>%
    kable_styling(
        latex_options = c("HOLD_position"),
        full_width = FALSE
    )
```

```{r save, include=FALSE}
source("io_checksum.R")
write_rds_w_checksum(df, "./data/dataset.rds")
```
