```{r, libraries_question, include=FALSE}
library(tidyverse)
library(cowplot)
library(corrplot)
```

# Questions

## Determine if the response variable (price) has an acceptably normal distribution. Address test to discard serial correlation.

```{r echo = FALSE}

df <- df %>% mutate(price_log = log(price))

price_mean <- mean(df$price)
price_sd <- sd(df$price)

ggplot(df, aes(price)) +
    geom_histogram(aes(y = ..density..), bins = 30) +
    stat_function(
        fun = dnorm,
        args = list(mean = price_mean, sd = price_sd),
        color = "red"
    ) +
    annotate("text", max(df$price) * 3 / 4, 1e-5,
        color = "red",
        label = sprintf(
            "μ = %0.2f\nσ = %0.2f",
            price_mean,
            price_sd
        ),
    )
```

```{r echo=FALSE, fig.height = 4}

price_log_mean <- mean(df$price_log)
price_log_sd <- sd(df$price_log)

ggplot(df, aes(price_log)) +
    geom_histogram(aes(y = ..density..), bins = 30) +
    stat_function(
        fun = dnorm,
        args = list(mean = price_log_mean, sd = price_log_sd),
        color = "red"
    ) +
    annotate("text", 11, 0.75,
        color = "red",
        label = sprintf(
            "μ = %0.2f\nσ = %0.2f",
            price_log_mean,
            price_log_sd
        ),
    )
```

```{r}
shapiro.test(df$price)
shapiro.test(log(df$price))
```

```{r echo=FALSE, fig.cap="\\label{fig:qq_plots}QQ plots", fig.height=3}
plt_qq_price <- ggplot(df, aes(sample = price)) +
    stat_qq() +
    stat_qq_line(color = "red") +
    theme(aspect.ratio = 1) +
    labs(x = "Theoretical", y = "price")
plt_qq_log_price <- ggplot(df, aes(sample = log(price))) +
    stat_qq() +
    stat_qq_line(color = "red") +
    theme(aspect.ratio = 1) +
    labs(x = "Theoretical", y = "log(price)")

plot_grid(plt_qq_price, plt_qq_log_price, align = "hv", labels = "AUTO")
```

```{r}
df$price %>% ks.test("pnorm", mean = mean(.), sd = sd(.))
df$price_log %>% ks.test("pnorm", mean = mean(.), sd = sd(.))
```

2. Indicate by exploration of the data, which are apparently the variables most
   associated with the response variable (use only the indicated variables). 

```{r, fig.cap = "Correlation plot"}
df %>%
    select(where(is.numeric)) %>%
    cor(use = "complete.obs") %>%
    corrplot()
```

3. Define a polytomic factor f.age for the covariate car age according to its
   quartiles, and argue if the average price depends on the level of age.
   Statistically justify the answer. 

```{r}
df %>%
    ggplot(aes(x = aux_age, y = price, fill = aux_age)) +
    geom_boxplot()

anova <- aov(price ~ aux_age, data = df)
summary(anova)
TukeyHSD(anova)
```


4. Calculate and interpret the anova model that explains car price according to
   the age factor and the fuel type. 

```{r}
df %>%
    ggplot(aes(x = aux_age, y = price, fill = fuelType)) +
    geom_boxplot()

anova <- aov(price ~ aux_age * fuelType, data = df)
summary(anova)

group_by(df, aux_age, fuelType) %>%
    summarise(
        count = n(),
        mean = mean(price, na.rm = TRUE),
        sd = sd(price, na.rm = TRUE)
    )
```

5. Do you think that the variability of the price depends on both factors? Does
   the relation between price and age factor depend on fuel type? 

6. Calculate the linear regression model that explains the price from the age:
   interpret the regression line and assess its quality.

7. What is the percentage of the price variability that is explained by the age
   of the car?

8. Do you think it is necessary to introduce a quadratic term in the equation
   that relates the price to its age? 

9. Are there any additional explanatory numeric variables needed to the car
   price? Study collinearity effects. 

10. After controlling by numerical variables, indicate whether the additive
    effect of the available factors on the price are statistically significant. 

11. Select the best model available so far. Interpret the equations that relate
    the explanatory variables to the answer (rate). 

12. Study the model that relates the logarithm of the price to the numerical
    variables. 

13. Once explanatory numerical variables are included in the model, are there
    any main effects from factors needed? 

14. Graphically assess the best model obtained so far. 

15. Assess the presence of outliers in the studentized residuals at a 99%
    confidence level. Indicate what those observations are. 

16. Study the presence of a priori influential data observations, indicating
    their number according to the criteria studied in class. 

17. Study the presence of a posteriori influential values, indicating the
    criteria studied in class and the actual atypical observations. 

18. Given a 5-year-old car, the rest of numerical variables on the mean and
    factors on the reference

19. Summarize what you have learned by working with this interesting real
    dataset.
